Vagueness, Noise, and Signaling

### 2015-01-18 paper/version_02/paper.tex
Moved diffusion as mutation to its own section

The next section introduces the background against which the work
presented here can be appreciated. Section~\ref{sec:repl-diff-dynam}
introduces the replicator diffusion dynamic and elaborates on its
relation with the replicator mutator
dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
diffusion dynamic. Section~\ref{sec:discussion} reflects and compares
our approach to others.

@@

The formulation of the \rdd given and explored here is in terms of
behavioral strategies, not in terms of mixed strategies (more on this
below). Dynamics on behavioral strategies assume that agents can adapt
their behavior locally, so to speak, i.e., independently at each
choice point. This greatly reduces the complexity of the dynamic and
simplifies numerical simulations. But it also points towards an
interpretation of the process as cultural evolution, rather than
biological evolution. Signaling strategies are not inherited as a
whole, but adapted piece-meal, e.g., by conditional imitation of
``local'' behavior at a given choice point (more below). Still, the
approach we suggest here is compatible with a formulation in terms of
mixed strategies and an interpretation as biological evolution as
well. To show this, we translate our initial formulation of the \rdd
into one using mixed strategies. Doing so identifies the \rdd as a
special case of the \rmd.

We begin by recapitulating the \rd in
Section~\ref{sec:repl-dynam-behav}, and introduce the \rdd in
Section~\ref{sec:repl-diff-dynam-1}.
Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
a special case of the \rmd. Sections~\ref{sec:repl-mutat-dynam} and
\ref{sec:diffusion-as-special} are conceptually important, but rather
technical and not strictly necessary to appreciate the main point of
the paper.

@@

The sequential definition highlights that other (discrete-time)
evolutionary dynamics can be subjected to state-confusability in an
analogous fashion. It would be entirely straightforward, for instance,
to perturb a best-response dynamic in case we wanted to model
cognizing, rational agents with a noisy perceptual system. Our choice
of the \rd as a carrier dynamic is motivated by its generality and
austerity, being compatible with a variety of implicit agent-level
behavior, without presupposing extravagant higher cognitive abilities
\citep[c.f.][]{Sandholm2013:Population-Game}. Moreover, by using the
\rd as a carrier for diffusion, we actually obtain a special case of a
behavioral-strategy version of the replicator mutator dynamics. This
is important to know because it enables similar applications of
diffusion from perceptual confusion also in models of biological
evolution. The remainder of this section therefore works out the
relation between the \rdd and the \rmd.

@@

\subsection{Diffusion as mutation}
\label{sec:diffusion-as-special}

@@

The previous section showed that the \rdd can be conceptualized as a
two-step procedure: we first calculate where the replicator dynamic
would take us, then apply diffusion. To understand the \rdd better, we
isolate the novel diffusion part in
Section~\ref{sec:iterated-diffusion}, before exploring how diffusion
and fitness-based selection interact in Section~\ref{sec:simulations}.

@@

The \rdd adds expected confusion of similar states to the \rd in
behavioral strategies. The result is an abstract evolutionary dynamic
that describes the most likely path of development of aggregate
population behavior in which any process of behavioral adaptation at
the agent level that is consistent with the \rd in behavioral
strategies is perturbed by systematic noise in the realization of the
agents' strategies. Given the use of behavioral strategies, an
interpretation as a process of cultural evolution is most salient,
according to which agents adapt their choice dispositions locally at
each choice point. But the \rdd is also extensible to biological
evolution on mixed strategies, in which case it is a special case of
the \rmd with mutation matrices derived from confusion matrices as
shown in Section~\ref{sec:diffusion-as-special}.

>>>>

The next section introduces the background against which the work
presented here can be appreciated. Section~\ref{sec:repl-diff-dynam}
introduces the replicator diffusion dynamic and Section~\ref{sec:diffusion-as-special}
elaborates on its relation with the replicator mutator
dynamic. Section~\ref{sec:exploring-rdd} explores the replicator
diffusion dynamic. Section~\ref{sec:discussion} reflects and compares
our approach to others.

@@

The formulation of the \rdd given and explored here is in terms of
behavioral strategies, not in terms of mixed strategies. Dynamics on
behavioral strategies assume that agents can adapt
their behavior locally, so to speak, i.e., independently at each
choice point. This greatly reduces the complexity of the dynamic and
simplifies numerical simulations. But it also points towards an
interpretation of the process as cultural evolution, rather than
biological evolution. Signaling strategies are not inherited as a
whole, but adapted piece-meal, e.g., by conditional imitation of
``local'' behavior at a given choice point (more below). Still, the
approach we suggest here is compatible with a formulation in terms of
mixed strategies and an interpretation as biological evolution as
well. In Section~\ref{sec:diffusion-as-special}, we
translate our initial formulation of the \rdd
into one using mixed strategies. Doing so identifies the \rdd as a
special case of the \rmd. For now, let us first focus on the
formulation of the dynamic in terms of behavioral strategies alone.

We begin by recapitulating the \rd in
Section~\ref{sec:repl-dynam-behav}, and introduce the \rdd in
Section~\ref{sec:repl-diff-dynam-1}.
%Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
%Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
%a special case of the \rmd. Sections~\ref{sec:repl-mutat-dynam} and
%\ref{sec:diffusion-as-special} are conceptually important, but rather
%technical and not strictly necessary to appreciate the main point of
%the paper.

@@

The sequential definition highlights that other (discrete-time)
evolutionary dynamics can be subjected to state-confusability in an
analogous fashion. It would be entirely straightforward, for instance,
to perturb a best-response dynamic in case we wanted to model
cognizing, rational agents with a noisy perceptual system. Our choice
of the \rd as a carrier dynamic is motivated by its generality and
austerity, being compatible with a variety of implicit agent-level
behavior, without presupposing extravagant higher cognitive abilities
\citep[c.f.][]{Sandholm2013:Population-Game}. Moreover, by using the
\rd as a carrier for diffusion, we actually obtain a special case of a
behavioral-strategy version of the replicator mutator dynamics. This
is important to know because it enables similar applications of
diffusion from perceptual confusion also in models of biological
evolution.

The following section works out the relation between the \rdd and the \rmd.
Although this section is conceptually important, it is also rather
technical and not strictly necessary to appreciate the main point of
the paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Diffusion as mutation}
\label{sec:diffusion-as-special}

@@

\subsection{Relating the dynamics}

@@

The previous sections showed that the \rdd can be conceptualized as a
two-step procedure: we first calculate where the replicator dynamic
would take us, then apply diffusion. To understand the \rdd better, we
isolate the novel diffusion part in
Section~\ref{sec:iterated-diffusion}, before exploring how diffusion
and fitness-based selection interact in Section~\ref{sec:simulations}.

@@

The \rdd adds expected confusion of similar states to the \rd in
behavioral strategies. The result is an abstract evolutionary dynamic
that describes the most likely path of development of aggregate
population behavior in which any process of behavioral adaptation at
the agent level that is consistent with the \rd in behavioral
strategies is perturbed by systematic noise in the realization of the
agents' strategies. Given the use of behavioral strategies, an
interpretation as a process of cultural evolution is most salient,
according to which agents adapt their choice dispositions locally at
each choice point. But the \rdd is also extensible to biological
evolution on mixed strategies, in which case it is a special case of
the \rmd with mutation matrices derived from confusion matrices, as
shown in Section~\ref{sec:diffusion-as-special}.
